---
title: "Assignment4: Data Visualization"
output: html_document
date: "2022-10-07"
---
## Group 3, Team 11:

* Robert Locher (5747465)
* Jose Enrique Leal Castillo (9066381)
* Niek Lieon (6520448)


# 1.- "Good" data visualization"

## Data Importing
```{r}
#Importing libraries to use
library(tidyverse)


#Reading dataset to use and saving it into a DataFrame variable
data <- read_csv('planets.csv', show_col_types = FALSE)

data

```

## Data Transformation
```{r}
#Creating density definition for planets
planet_dis <- data %>% 
  
  mutate(
    #Radius conversion from earth units to cm.
    new_pl_rade = pl_rade * 6.38E+08,
    #Mass conversion from earth units to gr.
    new_pl_bmasse = pl_bmasse * 5.97E+27
    ,
    #Density calculation (assumming sphere volume formula = 4/3 * Pi* r3)
    dens = new_pl_bmasse/((4/3)* 3.1416* (new_pl_rade*new_pl_rade*new_pl_rade))
    ,
    #Defining Rocky and Gas-Ice density threshold at 2.5 gr/cm3 (gassy and ice planets are oftenly less dense than rocky planets)
    #Assumption gathered from https://en.wikipedia.org/wiki/Terrestrial_planet
    planet_type = cut(dens, breaks = c(0, 2.5, 10000), 
                       labels = c("Gas-Ice","Rocky")))  %>% 

#Filtering na values
filter(!is.na(dens)) 

#Checking our results
select(planet_dis,new_pl_bmasse,new_pl_rade,planet_type,dens)

```
## Data Visualization
```{r}
#Defining planets graph
planets_graph <- 
  
  #Data to plot
  ggplot(planet_dis, aes(x =sy_dist , y =pl_rade, color = planet_type)) +
  #Scatterplot with alpha to see all points clearly
  geom_point(alpha = .7) +
  #Labels name
  labs(x = "Distance from Earth [P.C.]", y = "Planet size[U.E.]", color = "Planet type:")+
  #Straightening with log scales to avoid agglomeration
  scale_x_log10()+
  scale_y_log10()+
  #Adding a light time to increase data/ ink ratio
  theme_light()+
  #Nice descriptive title
  ggtitle("TESS discovered planets’ characteristics per year since 2018")+ 
  
  #Adding facet to include time factor in discoveries
  facet_wrap( ~ disc_year  ) 

#Plotting the final graph result  
planets_graph
```

## Why is this graph good?
This graph is good in our opinion because it represents a high quantity of data with low ink. 

We can observe a nice correlation between the distance from earth and the size of the planets, indicating us good insights about how this telescope detection method might work. As well, we introduce the time factor divided by year with the use of facets. Meanwhile, a useful color helps us  to visualize better which type of planet was discovered by the Tess telescope.

This plot uses a contrasting Hue to distinguish planet types which is one of the best ways to visualize categorical features according to Tamara Munzer. Finally, an important fact is that it uses the Tukey straightening approach in the scale of the axis to have a cleaner visualization of data.


# 2.- "Bad" data visualization"